% $Id$
\labeledchapter{Perl Features that make this possible}

We assume that you are already familiar with object oriented 
programming in Perl, or at least have some experience with it
in some other language, and that you have access to the Perl
documentation that comes with Perl.

All of the patterns that we discuss in this book are built
as modules, since part of the pattern design process is 
building reuseable components.  A lot of the Perl features
and techniques that we use to create a pattern revolve
around features for module creation or classes that are
typically used by modules, such as Exporter and AutoLoader.
You may already be familiar with the conventional ways of
using these features, but they have more power than that
if we think about them just a little differently than usual.
Indeed, one of the greatest compliments to any design is its
usefulness in ways never imagined.

    \section{How to use this chapter}

This chapter is a reference for the design patterns that we
discuss in the second half of this book.  If you are already 
familiar with object-oriented programming in Perl and the concepts
that you see in the section headings for this chapter then you
may want to skip ahead.  Otherwise, you may want to refer to this
chapter when you do not understand a technique or trick that we
use in a pattern.  If you are doing some sort of penance, then
you should read this entire chapter before going on to the next.

        \subsection{Use the Perl documentation.}

For more details on any of the concepts we discuss, you can get
more details by looking at the Perl documentation.  Every standard distribution
of Perl comes with documentation in a form appropriate for that
operating system.  You may also want to start using Perldoc.com which
has indexed and made searchable the Perl manual pages and a lot of the
documentation for popular modules.  If you have a web browser and access
to the internet then you can always get to the documentation.

Aside from the documentation, the definitive reference is {\it Programming Perl}.

\input{universal}

     \section{References}
     
     
     \subsection{Comparing references}

Have you ever accidently de-referenced an object, or even an ordinary
reference, while printing a string?  You probably got some funny looking
output that looked like the object's package name and what you might
want to use as a memory address.  Although there is no way in Perl to
take that stringified      

    \section{Inheritance}

Inheritance defines a familial relationship between classes, which may
have parents and children (or base and derived) classes.  Base classes
can define behavior for the objects that are derived from them, just
like all classes can use the behavior defined in UNIVERSAL.  Perl keeps
track of this using a special package array @ISA (pronounced ''is a'', as
in ''a camel is a mammal'').  Base class names are placed in the @ISA array.
Those base classes might also have their own base classes.  To complicate
matters even more, a class can inherit from more than one class at a time.
This leads to all sorts of problems unless we know that the multiple base
classes will not collide.  Several of the patterns use multiple inheritance,
but in a completely safe way.  The whole of the inheritance relationships
is a tree, with the UNIVERSAL class at its root.

\todo{nifty LaTeX figure of a tree}

When Perl cannot find a method in the derived class, it starts a depth first
"climb" (since it goes towards the root of the tree, which is actually
at the top) of the inheritance tree.  It first looks in the first class named 
in the current @ISA.  If it does not find the method in that class, it looks
in the first class named in that classes @ISA, so forth, working its way
back down and up if there is multiple inheritance.

\todo{example traversal}

    \section{Subclassing}

We can extend or modify a class by subclassing the base class.

    \subsection{SUPER}

The SUPER class is a special class name which connotes the immediate
parent class of the object.  It works on a per-object basis rather
than thinking about the current package or the package in which
it was defined.  This feature, which some people think is a 
mis-feature, is really quite handy for design patterns.  We can create
objects, re-bless them, pass them around, store them in complex structures,
and all other manner of chaos and SUPER still means the object's
parent class.

    \section{bless}

Perl is not an object oriented language, although it can do a very
good job pretending to be one.  Depending on your personal philosophy 
of language design, you probably either would give your right hand
for this or want to take Larry's right hand for making it that way.

Perl knows what class an instance thinks it is because we tag
the object with a string that is the class name.  That is really
all that bless does.  An object is really just a blessed reference,
or more simply, a reference that has a name tag.

\begin{quote}    
\begin{verbatim}
my $self = \%hash;

bless $self, $class
\end{verbatim}
\end{quote}

To read the name tag of an object, we can use ref() with which returns
the string bless() tagged it.  We could also use the UNIVERSAL::isa
function to check to see if the name tag is a particular string, as
we saw before, but we can also ref().

\begin{quote}    
\begin{verbatim}
my $scalar = 'foo';
bless $scalar, time;

my $time_created = ref $scalar;
\end{verbatim}
\end{quote}

There is absolutely no requirement that the class actually exist,
or that we use anything from the class with the tagged variable.
Indeed, one of the authors is perverse enough to use this tag
as out-of-band meta-information about the variable.  Consider
a variable

    \section{AutoLoader}

The most common use of Autoloader involves breaking up a module
with AutoSplit then loading methods only when they are needed. 
Large modules, such as \todo{some module names}, inherit from AutoLoader and put
their methods after the \_\_END\_\_ token.  When the module is
autosplit, each method is put into its own file in a special
directory.  At run-time, when an object calls a method which is
not known, the AUTOLOAD routine in the AutoLoader package is
called.  This routine looks into the special auto directory for
a file with the right name to match the method name and package
for the requested method.  If it is found it is loaded and the
program continues.

Autoloader only does this by default, however.  We can create our
own AUTOLOAD routine to do just about anything that we want.

One disadvantage of this approach is that there is presently no
way to re-enter the inheritance tree search process once we
have fallen into an AUTOLOAD routine in any of the packages we
were searching.  We either have to do something useful or give
up.  Rumor has it that the next version of Perl will be more
understanding.


    \section{Operator overloading}

We can change the way that operators behave with certain objects
by redefining what they do, or overloading them.  You probably
already recognize the common operators such as X, but may not
know about the operators XXX.  We can redefine the behavior
of each of these to suit our needs.


When we redefine an operator we change its behavior for all
instances in a class.  


    \section{Tied variables}

Perl has a nifty feature that allows us to use ordinary looking 
variables in the ordinary way but redefine what they actually do.
For instance, with a scalar variable we can assign to it, access
its value, interpolate it, and so on.  With tie we tell Perl that
these normal operations are actually going to be carried out by
a hidden object from a package that we create.  This package has 
special, predefined method names to carry out the operations of the
variable.



    \section{Comparison to other languages}

Despite what you may have been taught, there is a difference
between object oriented programming and object oriented languages.
This is not the typical assertion that an object oriented language
is a bondage-and-discipline language.  Indeed, C++ is a very strict 
language, but it is not an object oriented one even though you use
it to create classes and objects.  
 
        \subsection{C++}

C++ sucks for a number of reasons and we endeavour to list all of 
them here.

        \subsection{Java}

Java sucks less than C++, but suffers from an overzealous marketing
department.  

        \subsection{Python}

Python is a language for people who just want to do what they are
told and go home at the end of the day.        
        
        \subsection{Smalltalk}

Smalltalk is for geeks not satisfied with how things work and who
are brazen enough to change the class structure themselves.  They
distribute these changes as "Goodies" like some pervert giving 
tainted candy to halloween youngsters.

        \subsection{Eiffel}

Its named after something French.  I do not like France.

        \subsection{Ruby}

If you want to work in Perl for the rest of your life, you have to
kill Ruby which might be the next big thing.

        
